Characterization of the microbial and fungal communities in ancient pottery from Neolithic southwestern China

### 2024-03-03 analysis/paper/paper.qmd
add PCoA plots

The V6 region has similar bacterial communities at the phylum level as the V4 region across samples, with the exception of *Patescibacteria* (`r min(Pates$rel_ab)`- `r max(Pates$rel_ab)`%) in pottery samples, which was not detected in the reads of the V4 region. The dominant classes within *Patescibacteria* include *Parcubacteria*, *Microgenomatia*, and *Saccharimonadia*. We found a significant difference in the relative abundance of phyla between the pottery interior sample and soil samples from Gaoshan, but none for Guijiabao samples. At the family level, *Paenibacillaceae* was more abundant in pottery exterior samples than in pottery interior samples (P <0.05), while *Sphingomonas* was more abundant in pottery interior samples than in control samples. V1 and ITS1 regions have higher proportions of unidentified reads than V4 and V6 regions. The bacterial communities at the phylum level for the V1 region are similar to those for the V4 and V6 regions, although there were no significant variations in composition between samples. For the ITS1 region of 18S rRNA, fungal communities at the phylum level were predominated by *Ascomycota* (`r min(Ascom_GJB$rel_ab)`- `r max(Ascom_GJB$rel_ab)`% for Guijiabao samples and `r min(Ascom_GS$rel_ab)`- `r max(Ascom_GS$rel_ab)`% for Gaoshan samples) and *Basidiomycota* (`r min(Basid_GJB$rel_ab)`- `r max(Basid_GJB$rel_ab)`% for Guijiabao samples and `r min(Basid_GS$rel_ab)`- `r max(Basid_GS$rel_ab)`% for Gaoshan samples. 
```{r}
#| label: fig-V4-rel-ab
#| echo: false
#| cache: true
#| fig-cap: "A: The taxonomic composition of bacterial communities for the V4 region of pottery, soil, and control samples at the genus level. B: Bacterial communities at the family level differed significantly in relative abundance between pottery exterior and control samples, and between pottery exterior and interior samples."
library(here)
knitr::include_graphics(here("analysis", "figures", "V4_rel_sig_taxa_combine.png"))
```

## Beta diversity analysis

Alpha diversity 
Beta diversity
Relative abundance 
Venn diagrams
The abundant taxa at ASV level & the abundant taxa at genus level

>>>>

The V6 region has similar bacterial communities at the phylum level as the V4 region across samples, with the exception of *Patescibacteria* (`r min(Pates$rel_ab)`- `r max(Pates$rel_ab)`%) in pottery samples, which was not detected in the reads of the V4 region. The dominant classes within *Patescibacteria* include *Parcubacteria*, *Microgenomatia*, and *Saccharimonadia*. We found a significant difference in the relative abundance of phyla between the pottery interior sample and soil samples from Gaoshan, but none for Guijiabao samples. At the family level, *Paenibacillaceae* was more abundant in pottery exterior samples than in pottery interior samples (P <0.05), while *Sphingomonas* was more abundant in pottery interior samples than in control samples. V1 and ITS1 regions have higher proportions of unidentified reads than V4 and V6 regions. The bacterial communities at the phylum level for the V1 region are similar to those for the V4 and V6 regions, although there were no significant variations in composition between samples. For the ITS1 region of 18S rRNA, fungal communities at the phylum level were predominated by *Ascomycota* (`r min(Ascom_GJB$rel_ab)`- `r max(Ascom_GJB$rel_ab)`% for Guijiabao samples and `r min(Ascom_GS$rel_ab)`- `r max(Ascom_GS$rel_ab)`% for Gaoshan samples) and *Basidiomycota* (`r min(Basid_GJB$rel_ab)`- `r max(Basid_GJB$rel_ab)`% for Guijiabao samples and `r min(Basid_GS$rel_ab)`- `r max(Basid_GS$rel_ab)`% for Gaoshan samples.

```{r}
#| label: fig-V4-rel-ab
#| echo: false
#| cache: true
#| fig-cap: "A: The taxonomic composition of bacterial communities for the V4 region of pottery, soil, and control samples at the genus level. B: Bacterial communities at the family level differed significantly in relative abundance between pottery exterior and control samples, and between pottery exterior and interior samples."
library(here)
knitr::include_graphics(here("analysis", "figures", "V4_rel_sig_taxa_combine.png"))
```

## Beta diversity analysis
We used principal coordinates analysis (PCoA) based on binary Jaccard and Bray-Curtis dissimilarity to evaluate any structural differences between bacterial and fungal communities in four groups: control, pottery interior, pottery exterior, and soil samples.
The bacterial communities identified by the V1, V4, and V6 region sequences showed a significant difference between four sample groups for their presence and abundance according to PERMANOVA tests (P < 0.01), while there were no significant differences in the fungal communities (@fig-V4-V6-PCoA). A further pairwise PERMANOVA tests showed bacterial communities in pottery interior and exterior samples differed significantly from soil samples (P < 0.05) based on the V4 and V6 region sequences, but not on the V1 region sequences. Although there were no significant differences in beta diversity between pottery interior, pottery exterior, and control samples, the Jaccard PCoA plot (the V4 region) indicated a higher intra-variability between pottery samples from Guijiabao and Gaoshan, as well as soil samples from two sites according to the Jaccard PCoA plot (the V6 region) and the Bray-Curtis PCoA plots (the V4 and V6 regions).

```{r}
#| label: fig-V4-V6-PCoA
#| echo: false
#| cache: true
#| fig-cap: "A: PCoA plots of bacterial communities based on the sequences of V4 region. B: PCoA plots of bacterial communities based on the sequences of V6 region. The percentage variability explained by each principal component is indicated on the axes"
knitr::include_graphics(here("analysis", "figures", "V4_V6_combine_PcoA.png"))
```

Venn diagrams
The abundant taxa at ASV level & the abundant taxa at genus level
